---
title: "Exploratory Data Analysis"
author: "Jacob Gutierrez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: yeti
subtitle: "Thesis Project: SA2"
---

```{r knitr setup chunk, echo = F}
## Knitr setupchunk
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, rows.print=20, eval = TRUE)
```

# Purpose: {.tabset .tabset-fade .tabset-pills}

This script is expected to run on Hoolock2 See sessionInfo for more info.


# Analysis:

#### Libraries


```{r libraries}
library(tidyverse) ## Dataframe operations
library(methylKit) ## Methylation Sequencing tools
library(fastcluster) ## fast HC
library(uwot) ## UMAP
source("/home/groups/hoolock2/u0/jg/bin/FIt-SNE/fast_tsne.R") # fiT-SNE fftRtsne 
library(vizier) ## visualize embedings ## Use devtools ot install may replace with custom.
library(plot3D)
source('../scripts/Methylation_SA2/mc_analysis_tools.R')  # SVA code and PC-PR2 code and PCA code
source('../bin/jg_functions.R') # Pretty html printing. 
```


### Data read in 

```{r mapping file readin}
exp_name ="ECP15"

#c_mappings <- read.csv('ecp15_mapping.csv',stringsAsFactors = FALSE)
c_mappings <- readxl::read_xlsx(path="/home/groups/hoolock2/u0/jg/thesis/mcseq_data/mc-seq_mapping_plus15.xlsx", sheet = exp_name)


# Convering to characters as methykit is picky Done above
c_mappings <-  c_mappings %>% mutate(complex = case_when( Karyotype == '45,X' ~ 0,
                                                          Karyotype == 'N/A' ~ 0, 
                                                          TRUE ~ 1))
## Make factors
c_mappings <- c_mappings %>% mutate(BAV = as.factor(BAV), Smoking = as.factor(Smoking), complex = as.factor(complex), Pool = as.factor(Pool))


## Add text_bav 
c_mappings <- c_mappings %>% mutate(bav_tav = as.factor(if_else(BAV == 1, "BAV","TAV")))

c_mappings %>% jg_pretty_print()
```

> Read in Imputed Matrix



```{r read in methylation}
imputed_mat <-  readRDS('meth_data/tiled_before_imputed_meth.rds')
clean_mat <- remove_invariance(imputed_mat)
#most_var_mat <- 
```

## Heirarchical Clustering

I will use the fast cluster package as there is a known bug in the stats::hclust() function. 

Use correlation distance! 

```{r Correlation Function}
dist.cor <- function(x, method = "pearson", abs=TRUE, diag=FALSE, upper=FALSE){
  
  xcor <- cor(t(x), method = method)
  
  if(abs){
    xcor = 1-abs(xcor)
  } else{
    xcor = 1-xcor
  }
  
  d <- xcor[lower.tri(xcor,diag=FALSE)]
  
  
  attr(d, "Size") <- nrow(x)
  attr(d, "Labels") <- dimnames(x)[[1L]]
  attr(d, "Diag") <- diag
  attr(d, "Upper") <- upper
  attr(d, "method") <- method
  attr(d, "call") <- match.call()
  class(d) <- "dist"
  
  return(d)
}
```





```{r Clustering Plots}
#dist_mat <- dist(scale(t(clean_mat)))
cor_dist <- dist.cor(t(clean_mat))
hc_out <- hclust(cor_dist,method = "complete")
plot(hc_out)

```


```{r Other Clustering Plots}
new_hc <- hc_out
new_hc$labels <- c_mappings$BAV
plot(new_hc)
```


## PCA

```{r Run PCA}
batch_factors <- c("BAV", "Age","Smoking","complex","Pool")
atitle <- "Preliminary Methyl-Capture Data"
mc_pca <- assess_pca(mat = clean_mat,pd = c_mappings, batch.factors = batch_factors, title = atitle)
mc_pca$heatmap
```

> Need to fix the color scaling. 

## PC-PR2



```{r PCPR2 Run}
tit = "PC-PR2 Analysis of Known Sources of Variation"
PC_PR2(mc_pca$pca, c_mappings, batch_factors,tit)
```

## UMAP

umap(
  X,
  n_neighbors = 15,
  n_components = 2,
  metric = "euclidean",
  n_epochs = NULL,
  learning_rate = 1,
  scale = FALSE,
  init = "spectral",
  init_sdev = NULL,
  spread = 1,
  min_dist = 0.01,
  set_op_mix_ratio = 1,
  local_connectivity = 1,
  bandwidth = 1,
  repulsion_strength = 1,
  negative_sample_rate = 5,
  a = NULL,
  b = NULL,
  nn_method = NULL,
  n_trees = 50,
  search_k = 2 * n_neighbors * n_trees,
  approx_pow = FALSE,
  y = NULL,
  target_n_neighbors = n_neighbors,
  target_metric = "euclidean",
  target_weight = 0.5,
  pca = NULL,
  pca_center = TRUE,
  pcg_rand = TRUE,
  fast_sgd = FALSE,
  ret_model = FALSE,
  ret_nn = FALSE,
  n_threads = max(1, RcppParallel::defaultNumThreads()/2),
  n_sgd_threads = 0,
  grain_size = 1,
  tmpdir = tempdir(),
  verbose = getOption("verbose", TRUE)
)

n_sgd_threads >1 RESULTS NOT REPRODUCIBLE! 


```{r UMAP}
## JG 8/19/20: Umap without group seeding. 
## Checking for files:
prev_umap <- "eda_data/umap_embeddings.rds"
if (file.exists(prev_umap)){
  ## THe file exists load it.
  print("loading Previously Generated UMAP Object!")
  imput_umap <- readRDS(prev_umap)
}else{
  # It doesnt exist so run it 
  ## JG ADD SAVING IT LATER!!!!!
  print("No Output Detected Running UMAP")
  imput_umap <- umap(t(clean_mat),init = 'spca',n_components = 3, n_neighbors = 10, a = 2, b = 2, local_connectivity = 2, fast_sgd = TRUE)
  
  ## Saving
  saveRDS(imput_umap, prev_umap)
}

## JG 8/19/20: Umap without group seeding. 
# imput_umap <- umap(t(clean_mat),init = 'spca',n_components = 3, n_neighbors = 10, a = 2, b = 2, local_connectivity = 2, fast_sgd = TRUE)


## JG 8/19/20: Umap WITH group seeding. 
#imput_umap <- umap(t(clean_mat),init = 'spca', n_neighbors = 10, a = 2, b = 2, local_connectivity = 2, fast_sgd = TRUE, y = c_mappings$BAV)
#saveRDS(imput_umap,'T_Umap_saved.RDS')
#saveRDS(object = imput_umap, file = "data/umap_embeddings.rds")
#imput_umap <- readRDS("data/umap_embeddings.rds")

the_title <- "UMAP visualization"
embed_plot(imput_umap, c_mappings$BAV,text = c_mappings$bav_tav, title = the_title)
## 

```



```{r UMAP 3D}
scatter3D(imput_umap[,1], imput_umap[,2], imput_umap[,3], colvar = NULL, col = c_mappings$bav_tav,
          pch = 19, cex = 0.5)
```


> Make a function to explore all these options and save the params into the plot! Then make all of them and write to pdf

> BAV appears to be seperate from TAV because I seeded the algorithm.

## T-SNE


> same as above!

```{r Run TSNE}

## Checking for files:
prev_tsne <- "eda_data/tsne_embeddings.rds"
if (file.exists(prev_tsne)){
  ## THe file exists load it.
  print("loading Previously Generated TSNE Object!")
  tnse_out <- readRDS(prev_tsne)
}else{
  # It doesnt exist so run it 
  ## JG ADD SAVING IT LATER!!!!!
  print("No Output Detected Running TSNE")
  tnse_out <- fftRtsne(t(clean_mat), fast_tsne_path = "/home/groups/hoolock2/u0/jg/bin/FIt-SNE/bin/fast_tsne", dims = 2, perplexity = 1, max_iter = 1000)
  
  ## Saving
  saveRDS(tnse_out, prev_tsne)
}
# tnse_out <- fftRtsne(t(clean_mat), fast_tsne_path = "/home/groups/hoolock2/u0/jg/bin/FIt-SNE/bin/fast_tsne", dims = 2, perplexity = 1, max_iter = 1000)
# 
# saveRDS(tnse_out, "eda_data/tsne_embeddings.rds")

```


Visualization Here

```{r TSNE Vis}
the_title <- "Tsne visualization"
embed_plot(tnse_out, c_mappings$BAV,text = c_mappings$bav_tav, title = the_title)
```



## SVA


```{r SVA Params}
variable_imputed <- remove_invariance(clean_mat)
all_covariates <- c('bav_tav', 'Age','Smoking','complex','Pool')
main_var <- 'bav_tav'
biological <- c('Age')
naive_title <- 'Value before imputation'
```

> It appears 8% of all CpGs are the same across all samples and were removed for cell type composition estimation. 

> Now applying the Traditional SVA algorithm 

```{r RUN SVA}
## Estimating the Number of SVs by permutation 
## Checking for files:
prev_sva <- "eda_data/sva_out.rds"
if (file.exists(prev_sva)){
  ## THe file exists load it.
  print("loading Previously Generated SVA Object!")
  tile_sva <- readRDS(prev_sva)
}else{
  # It doesnt exist so run it 
  ## JG ADD SAVING IT LATER!!!!!
  print("No Output Detected Running SVA")
  tile_sva <- sva_assessment(variable_imputed, c_mappings, all_covariates, main_var, biological, naive_title, discover = FALSE, algorithm = "Leek", sv_est_alg = 'be') #'be') #"RMT")
  
  ## Saving
  saveRDS(tile_sva, prev_sva)
}

tile_sva$heatplot

tile_sva$barplot

SV <- tile_sva$sv
colnames(SV) <-  paste("SV", seq(tile_sva$n.sv), sep ='')

```

<!-- Saving resulting surrogate Varibles -->

<!-- ```{r} -->
<!-- saveRDS(tile_sva,'data/sva_out.rds') -->


<!-- ``` -->




## SVA PC-PR2

> NULL PCPR2

```{r sva PCPR2}
pheno <- c_mappings[batch_factors]


tit = "PC-PR2 Analysis of Known Sources of Variation"
PC_PR2(mc_pca$pca, pheno, batch_factors,tit)
```


> Adding SVs to phenotype data> 


```{r PCPR2 SV + Bio}
bio <- pheno[,1:2] 
bio <- cbind(bio, SV)

tit = "PC-PR2 Analysis of Biological Factors & Surrogate Variables"
PC_PR2(mc_pca$pca, bio, colnames(bio),tit)
```

> Now Assessing everything



```{r SVA ALL}
pheno <- cbind(pheno, SV)

tit = "PC-PR2 Analysis of All Sources of Variation"
PC_PR2(mc_pca$pca,pheno , colnames(pheno),tit)
```

> Assessing Just SVA



```{r PCPR2 Just SVA}
pheno <- cbind(pheno, SV)
col_vec <- colnames(pheno)
sv_name <- col_vec[col_vec %>% grep("SV.",.)]

tit  <-  "PC-PR2 Analysis of All Sources of Variation"

PC_PR2(mc_pca$pca,pheno , sv_name,tit)
```

In the preliminary analysis Surrogate variables explain rough 37% of the data. 



### Packages

```{r R info}
sessionInfo()
```

