---
title: "BLANK"
author: "Jacob Gutierrez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: yeti
subtitle: "Thesis Project: SA1"
---

```{r knitr setup chunk, echo = F}
## Knitr setupchunk
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, rows.print=20, eval = TRUE)
```

# Purpose: {.tabset .tabset-fade .tabset-pills}

This script is expected to run on Hoolock2 See sessionInfo for more info.

Report will render reports and execute code related to the Analysis from **Raw** Bismark Coverage reports to EDA plots and DMR annotation info. I can also add the results of the analysis in this summary report. 




# Run Analysis:

## Methylation QC

```{bash make report directory, eval = F}
mkdir -p meth_qc
```


```{R Render Report , eval = F}
outfile <- sprintf("meth_qc/%s_Methylation_QC", "ECP15")
rmarkdown::render("meth_sample_qc_template.Rmd", output_file = outfile)
```


### Results: 

Not sure what Figures or Dataframes to export....

* Number of CpGs to learn on.
* Number of Tiles to learn on
* Number of Methylation Values Imputed


## EDA Report



Analyses done: 
1) Clustering 
2)PCA
3) PC-PR2
4) SVA
5) UMAP
6) T-SNE

### Results

If you would like to see all results please explore the html report.
Place EDA highlights here...

* SVA summary
* Best of plots. 

Not sure what Figures or Dataframes to export....

## Differential Methylation & DMR Annotation

Rendering Diff Meth Report

Rendering Annotation Report

### Results


* Number of Significant CpG
* Number of Significant Tiles

Not sure what Figures or Dataframes to export....


# Biological Inference

Code to do Downstream Analysis Here

## Enrichment Analysis

Render Report Here
```{r}

```


Highlight Results here

## String Database


## GREAT Analysis

## TF Motif Enrichment


Run Homer Script Here


# Full Results Summary

Complete Results Highlights.



### Packages

```{r R info}
sessionInfo()
```

