---
title: "BLANK"
author: "Jacob Gutierrez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: yeti
subtitle: "Thesis Project: SA1"
---

```{r, echo = F}
## Knitr setupchunk
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, rows.print=20, eval = TRUE)
```

# Purpose: {.tabset .tabset-fade .tabset-pills}

This script is expected to run on Hoolock2 See sessionInfo for more info.

Here I will take the lifted-over manifest file and create a track. 


# Analysis:

#### Libraries


```{r libraries}
library(rtracklayer) ## Bed tool package
```


### Data read in 

Read in data 
```{r}
hg38_gr <- import.bed("hg38-ensemble-truseq-manifest.bed")
```

## Create manifest

Place annotation into UCSC format for genome browser. 

```{r}
seqlevelsStyle(hg38_gr) <- "UCSC"
```


Removing all non canonical chromosomes

```{r}
canonical <- paste("chr",c(seq(22),"X","Y"), sep = "")
hg38_gr <- keepSeqlevels(hg38_gr,canonical, pruning.mode = "coarse")
```

Write out 

```{r}
export.bed(hg38_gr, "ucsc_format.bed")
```

> END OF R CODE

### BASH Processing

Here I will just combine the previously made bed file with the header. 

```{bash}
cat manifest_track_header.txt ucsc_format.bed > hg38-mcseq-truseq-track.bed

gzip hg38-mcseq-truseq-track.bed

## Clear buffer
rm ucsc_format.bed
```





### Packages

```{r}
sessionInfo()
```

